---
title: "Comparision of Non-nest Models"
author: "Edgar C. Merkle and Dongjun You"
date: "`r Sys.Date()`"
output: 
  rmarkdown::pdf_document:
    includes:
      in_header: nonnest2-preamble.tex
vignette: >
  %\VignetteIndexEntry{Comparision of Non-nest Models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{nonnest2}
---

# An Introduction

```{r echo=FALSE, message=FALSE}
## global knitr options
knitr::opts_chunk$set(fig.path='figure/nonnest2-', fig.align='center',
                      fig.show='hold', size='footnotesize', 
                      cache.path="cache/", warning=FALSE, message=FALSE)

## packages
library("simsem")
library("nonnest2")
```

## 1. Installation

You can install **nonnest2** from [CRAN](http://cran.rstudio.com/package=nonnest2), if you want to test the latest development version:

```{r eval=FALSE}
install.packages('nonnest2')

# to install the development version, run
# library(devtools)
# install_github("qpsy/nonnest2")
```


# 2. Analysis

## 2.1. Non-distinguishable Models

\begin{figure}[H]
\centering
\resizebox{0.45\textwidth}{!}{
\input{diagram}
}
\caption{Population Model}
\end{figure}

\begin{figure}%
  \centering
  \begin{subfigure}[t]{0.5\textwidth}%
    \centering
    \captionsetup{justification=centering,font=scriptsize}
    \resizebox{0.9\textwidth}{!}{
      \input{diagramA}
    }
    \caption*{Model A}\label{fig:modelA}
  \end{subfigure}%
  ~
  \begin{subfigure}[t]{0.5\textwidth}%
    \centering
    \captionsetup{justification=centering,font=scriptsize}
    \resizebox{0.9\textwidth}{!}{
      \input{diagramB}
    }
    \caption*{Model B ($\xi_2 \rightarrow X_3$)}\label{fig:modelB}
  \end{subfigure}%
  \caption{Path diagrams reflecting the two competing models}
  \label{fig:competingModels}
\end{figure}

```{r model_pop1, echo=FALSE}
# population model
loading <- matrix(0, 6, 2) 
loading[1:3, 1] <- NA 
loading[4:6, 2] <- NA
# population values
loadingValues <- matrix(0, 6, 2)  
loadingValues[1:3, 1] <- 0.7  
loadingValues[4:6, 2] <- 0.7
# factor loading object
LX <- bind(loading, loadingValues)

# error covariances
errorCor <- diag(nrow=6, ncol=6)
RTE <- binds(errorCor)

# latent covariances
latentCor <- matrix(NA, 2, 2)
diag(latentCor) <- 1
RPS <- binds(latentCor, 0.5)

modelpop1 <- model(LY = LX, RPS = RPS, RTE = RTE, modelType="CFA")
set.seed(2015)
dat1 <- generate(modelpop1, 200)
```

```{r model_A, echo=FALSE}
# model A
loadingA <- matrix(0, 6, 2) 
loadingA[1:4, 1] <- NA # 1:4 
loadingA[4:6, 2] <- NA

# factor loading object
LXA <- bind(loadingA)

errorCorA <- diag(NA, nrow=6, ncol=6)
RTEA <- binds(errorCorA)

latentCorA <- matrix(NA, 2, 2)
RPSA <- binds(latentCorA)
modelA_gen <- model(LY = LXA, RPS = RPSA, RTE = RTEA, modelType="CFA")
modelA <- modelA_gen@pt
```

```{r model_B, echo=FALSE}
# model B
loadingB <- matrix(0, 6, 2) 
loadingB[1:3, 1] <- NA # 1:4 
loadingB[3:6, 2] <- NA

# factor loading object
LXB <- bind(loadingB)

errorCorB <- diag(NA, nrow=6, ncol=6)
RTEB <- binds(errorCorB)

latentCorB <- matrix(NA, 2, 2)
RPSB <- binds(latentCorB)
modelB_gen <- model(LY = LXB, RPS = RPSB, RTE = RTEB, modelType="CFA")
modelB <- modelB_gen@pt
```

```{r echo=TRUE}
modelA_1 <- cfa(modelA, data=dat1)
modelA_1

modelB_1 <- cfa(modelB, data=dat1)
modelB_1
```

```{r}
vuongtest(modelA_1, modelB_1)
```

```{r}
icci(modelA_1, modelB_1)
```

## 2.2. Distiguishable and Non-Nested models

\begin{figure}[H]
\centering
\resizebox{0.45\textwidth}{!}{
\input{diagramA}
}
\caption{Population Model}
\end{figure}

```{r model_pop2, echo=FALSE}
# population model
loading2 <- matrix(0, 6, 2) 
loading2[1:4, 1] <- NA 
loading2[4:6, 2] <- NA
# population values
loadingValues2 <- matrix(0, 6, 2)  
loadingValues2[1:3, 1] <- 0.7
# reduce the value to prevent negative variance
loadingValues2[4, 1] <- 0.3
loadingValues2[4:6, 2] <- 0.7
# factor loading object
LX2 <- bind(loading2, loadingValues2)

modelpop2 <- model(LY = LX2, RPS = RPS, RTE = RTE, modelType="CFA")
set.seed(2015)
dat2 <- generate(modelpop2, 200)
```

```{r echo=TRUE}
modelA_2 <- cfa(modelA, data=dat2)
modelA_2

modelB_2 <- cfa(modelB, data=dat2)
modelB_2
```

```{r}
vuongtest(modelA_2, modelB_2)
```

```{r}
icci(modelA_2, modelB_2)
```